Competing Bandits: \\ The Perils of Exploration under Competition

### 2018-11-04 paper/content/introduction.tex
updated intro

\gaedit{We ask whether algorithms that are traditionally considered ``better" in the standard multi-armed bandit problem are incentivized to be deployed under competition.} We investigate this issue via extensive numerical experiments in a stylized duopoly model. In our model, two firms compete for users \swdelete{on a single dimension, quality},\swcomment{not sure what this means} and simultaneously learn from them. Each firm commits to a multi-armed bandit algorithm, and \emph{explores} according to this algorithm. Users select between the two firms based on the current reputation score: rewards from the firm's algorithm, averaged over a recent time window. Each firm's objective is to maximize its \gaedit{expected} market share (the fraction of users choosing this firm).

@@

\swdelete{Even though, under our model, consumers are best off when there is a ``temporary" monopoly, a recent subject of debate in policy circles is whether the data advantage alone poses a sufficiently powerful barrier to entry to warrant anti-trust interest.}
\swedit{Beyond the welfare implications, we also investigate the barrier of entry from the incumbant's data advantage.}
 While \cite{varian2018artificial} and \cite{bajari2018impact} argue that, from a statistical perspective, data has decreasing returns to scale and therefore does not serve as a significant competitive advantage, they focus on the problem in isolation. We argue that, to reason about this question, one needs to consider what the impact of better data is when data collection has competitive consequences, especially when firms need to learn while competing. Therefore, we run additional experiments that attempt to isolate the impact of the data advantage and reputation advantage. We find that the data advantage alone still leads to a significant incumbency advantage. Additionally, the data advantage tends to be stronger when the incumbent commits to a more advanced bandit algorithm. This highlights that, when thinking about data as a barrier to entry, data quality and not just quantity matters. In particular, the data advantage may be larger and more relevant when the incumbent possesses data that it gets via exploration of potentially sub-optimal actions that would be costly for the entrant to gather upon entry.}\swcomment{guy, could you make another pass over above}

>>>>

\gaedit{We ask whether algorithms that are traditionally considered ``better" in the standard multi-armed bandit problem are incentivized to be deployed under competition.} We investigate this issue via extensive numerical experiments in a stylized duopoly model. In our model, two firms compete for users and simultaneously learn from them. Each firm commits to a multi-armed bandit algorithm, and \emph{explores} according to this algorithm. Users select between the two firms based on the current reputation score: rewards from the firm's algorithm, averaged over a recent time window. Each firm's objective is to maximize its \gaedit{expected} market share (the fraction of users choosing this firm).

@@

\swedit{Beyond the welfare implications, we also investigate the barrier to entry that follows from the incumbent's data advantage.}  We run additional experiments that attempt to isolate the impact of the data advantage and reputation advantage. We find that the data advantage alone still leads to a significant incumbency advantage. Additionally, the data advantage tends to be stronger when the incumbent commits to a more advanced bandit algorithm. Our results highlight the importance of considering competition dynamics when thinking about the anti-trust question of whether data can serve as a significant barrier to entry in an industry. While data has decreasing returns to scale from a statistical perspective, in a competitive environment the relative differences in the data possessed by the firms may lead to a large competitive benefit due to the feedback loops discussed previously. In our model this effect is amplified when the incumbent commits to a more advanced bandit algorithm since, in the incumbency period, the firm gathers data on possible actions more efficiently leading to the relative gap upon entry to be larger. This is important when exploration can have larger reputational consequences.}\swcomment{guy, could you make another pass over above}
